---
title: "Introduction to unga"
author: "Peter Thuresson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to unga}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The main purpose with this package is to analyze data from a questionnaire.
You will be able to:

1) make contingency tables of every variable in a data frame.
2) cross tabulate all variables against an explanatory variable.
3) summarize some of the response alternatives of questions in a questionnaire.
4) overview the data and the dimensions of all questions, i.e. explore how 
many response alternatives every question has in a data frame.
5) find the item non response levels.

<h2>Let's begin!</h2>


```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(unga)
set.seed(1014)
```

First use library 'unga' to load functions.

```{r setup}
library(unga)

```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<h2>Overview data</h2>

Assume that you want to get a quick overview of your data frame with respect to 
how many response alternatives every question have in a survey. Let's use the 
function `table_alt()`. This is a good way to start when you want to use 
`table_unga()` and extract questions with a specific number of response 
alternatives. Let's generate some test data and apply the function `table_alt()`
to it.

```{r data2}
set.seed(123456)
gender<-round(runif(50,1,2))
variable_1<-round(runif(50,1,2))
variable_2<-round(runif(50,1,5))
variable_3<-round(runif(50,1,5))
df<-data.frame(gender,variable_1,variable_2,variable_3)

table_alt(df)

```

The output here tells you that you have two questions with two response
alternatives, and two questions with 5 response alternatives. 

<h2>Specific cross tabulation</h2>

Let's now apply the function `table_unga()` on questions with 5 response
alternatives on it and use `gender` as our explanatory variable.

```{r data 5}
table_unga(5,1,gender,df)
```

What we have done is to apply the function `table_unga()` to all variables in 
the data frame.
We have selected all (hypothetical) questions with 5 response alternatives (the
first argument in the function) in the data frame and used "gender" as our 
explanatory variable (the third argument in the function). The output here is 
proportion tables of those questions. The fourth argument is the data frame.

As described above, the first argument in `table_unga()` let you extract 
questions with a specified number of response alternatives. The third argument 
is  the explanatory variable, and the fourth argument is the data frame. 
But what about the second argument?
You could easily ignore this if you want. But it has to be specified to 1 or 2.
This will not affect the output in `Output_1`, i.e. what is printed by
`table_unga()` function.
The second argument let you summarize response alternatives, but only for
questions with 5 or 6 response alternatives. The output is found in `Output_2`.
Let us again look at our example but now look on `Output_2`:


```{r data3}

Output_2

```


This means that you have summarized response alternatives for questions with
5 response alternatives so now you have only 6 rows instead of 10 rows. 
The second argument `1` in the function means that you summarize alternatives 
according to: `(A+B,C+D,E)`. If you use `2` as input in the second argument you 
will summarize alternatives according to: `(A+B,C,D+E)`. 

You can also summarize questions with 6 response alternatives.
The summation will be done according to: `(A+B,C,D+E,F)`. Make sure you specify
your secound argument to `1`. The summarized output will again be found in
`Output_2`.

<h2>General cross tabulation</h2>

If you want to cross tabulate on all variables in your data frame without 
extracting questions with a specific response alternative, use 
`table_unga_general()`. This will return a list of contingency tables.

<h2>Item non response</h2>

The function `table_non_resp()` let you overview item non response within a data
frame. 

```{r data4}

set.seed(123456)
gender<-round(runif(50,1,2))
variable_1<-sample(c(1,2,3,4,5,6,NA),50,replace = TRUE)
variable_2<-sample(c(1,2,3,4,5,6,NA),50,replace = TRUE)
variable_3<-sample(c(1,2,3,4,5,6,NA),50,replace = TRUE)
df<-data.frame(gender,variable_1,variable_2,variable_3)

table_non_resp(df)

```


The output shows for example that `variable_1` is missing 10 percent of items.
If you want the absolute values, use the function `table_non_resp_2()` 
